**Density Dependence**

### 2023-02-09 Rmd/ddMac.Rmd
checked

---
title: "**Density Dependence**"
subtitle: "Modelling density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE

@@

Mackerel life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age. The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.

@@

$W_a = {\alpha}_a B^{{\beta}_a}$

Where $\alpha$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data. $W_a$ is constrained so sizes do not become unfeasible large or small.

@@

Maturity is modelled as a logistic function of mass-at-age i.e.

@@

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/dd/",
               fig.path  ="../tex/figs/dd/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

@@

```{r}
source("../R/dd.R")
```

@@

```{r, mac.data}
load(file.path(dropboxdir,"data/om/mac.RData"))
ts =as.data.frame(read_excel("../data/inputs/dat.xlsx",3))
```

```{r eq}
par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
par=lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)
eq =lhEql(par,m="lorenzen")

p1=plot(eq,refpt="msy")
```

```{r}
om=as(eq,"FLStock")
om=fwd(om,fbar=fbar(om)[,-1],sr=eq)

p2=plot(om)
```

```{r, eval=FALSE}
plot(qmin(qmax(0.96*biomass(eq)%/%biomass(eq)[,14],0.5),1.5))+
  scale_y_continuous(labels = scales::percent)+
  geom_vline(aes(xintercept=14),col="red")+
  xlab("Biomass")+ylab("% Change in mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Density dependence multiplier, red is $B_{MSY}$

```{r}
save(eq,file="~/Desktop/eq.RData")
```

```{r}
wt=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"]) 
ggplot(as.data.frame(wt))+           
    geom_line(aes(age,data,col=factor(year)))+
    theme(legend.position="none")+
    scale_x_continuous(limits=c(0,15))+
    xlab("Age")+ylab("Mass-at-age") 
```

@@

```{r}
parMat=FLPar(c(k=0.1, w5=111.0))
mat=1/(1+exp(-parMat[1]*(wt-parMat[2])))

ggplot(as.data.frame(mat))+           
    geom_line(aes(age,data,col=factor(year)))+
    theme(legend.position="none")+
    scale_x_continuous(limits=c(0,15))+
    xlab("Age")+ylab("M-at-age")
```

@@

```{r}
ggplot(ddM(stock.wt(eq),fbar(eq)))+
    geom_line(aes(age,data,col=factor(year)))+
    theme(legend.position="none")+
    scale_x_continuous(limits=c(0,15))+
    xlab("Age")+ylab("M-at-age") 
```

**Figure `r iFig=iFig+1; iFig`** Pope: simulation of density dependence in M-at-age

```{r}
m=par["m1"]%*%(wt%^%par["m2"])

ggplot(as.data.frame(m))+           
    geom_line(aes(age,data,col=factor(year)))+
    theme(legend.position="none")+
    scale_x_continuous(limits=c(0,15))+
    xlab("Age")+ylab("M-at-age") 
```

@@

```{r}
save(eq,file="/home/laurie/Desktop/tmp/t.RData")
```

@@

```{r, dd-eq-mass-mat-m}
x=propagate(eq,length(dimnames(eq)$year))
fbar(x)[,1]=c(seq(0,max(fbar(eq)),length.out=51),seq(1,50,length.out=51)*max(fbar(eq)))[-52]
fbar(x)=fbar(x)[,1]

@@

```{r}
save(eq,parMat,par,sr,file="/home/laurie/Desktop/tmp/t.RData")
```

```{r}
load("/home/laurie/Desktop/tmp/t.RData")

par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","ssb"]),b=0.3,k=0.1, w50=111.0))

@@

om2=om
lsr=landings.wt(om2)/stock.wt(om2)
dsr=discards.wt(om2)/stock.wt(om2)

@@

plot(window(FLStocks("OM"=om,DD=om2),start=11))
```

**Figure `r iFig=iFig+1; iFig`** Mass, maturity and M-at-age: Comparison of projections with and without density dependence.

>>>>

---
title: "**Density Dependence**"
subtitle: "Mackerel: density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE

@@

Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age. The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.

@@

$W_a = {\overline{W_a}} B^{{\beta}_a}$

Where $\overline{W_a}$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data. $W_a$ is constrained so sizes do not become unfeasible large or small.

@@

Maturity is then  modelled as a logistic function of mass-at-age i.e.

@@

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/mac/dd/",
               fig.path  ="../tex/figs/mac/dd/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

@@

```{r, cache=FALSE}
source("../R/dd.R")
```

@@

```{r, eq}
par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
par=lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)
eq =lhEql(par,m="lorenzen")
```

```{r, stk}
om=as(eq,"FLStock")
om=fwd(om,fbar=fbar(om)[,-1],sr=eq)
```

```{r, prdFn}
ggplot(model.frame(FLQuants(Biomass=biomass(eq),Yield=catch(eq))))+
  geom_line(aes(Biomass,Yield))
```

**Figure `r iFig=iFig+1; iFig`** Yield v Biomass at equilibrium

```{r}
dat=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"]) 
dat=merge(as.data.frame(dat,drop=T),as.data.frame(biomass(eq)%/%refpts(eq)["msy","biomass"],drop=T),by="year")
names(dat)[3:4]=c("mass","biomass")

ggplot(dat)+           
    geom_line(aes(age,mass,col=biomass,group=year))+
    scale_x_continuous(limits=c(0,6))+
    scale_y_continuous(limits=c(0,700))+
    geom_line(aes(age,data),data=as.data.frame(stock.wt(eq)),col="red",size=2)+
    xlab("Age")+ylab("Mass-at-age")+
    theme(legend.position="none")
```

@@

```{r}
parMat=FLPar(c(k=0.1, w5=par["a"]*par["l50"]^par["b"]))
dat=transform(dat,mat=1/(1+exp(-parMat[1]*(mass-parMat[2]))))

ggplot(dat)+           
    geom_line(aes(age,mat,col=biomass,group=year))+
    scale_x_continuous(limits=c(0,6))+
    geom_line(aes(age,data),data=as.data.frame(mat(eq)),col="red",size=2)+
    xlab("Age")+ylab("Proportion mature-at-age")+
    theme(legend.position="none")
```

@@

```{r}
dat=transform(dat,m=c(par["m1"])*(mass^c(par["m2"])))

ggplot(dat)+           
    geom_line(aes(age,m,col=biomass,group=year))+
    scale_x_continuous(limits=c(0,15))+
    geom_line(aes(age,data),data=as.data.frame(m(eq)),col="red",size=2)+
    theme(legend.position="none")+
    xlab("Age")+ylab("M-at-age") 
```

@@

@@

```{r, dd-eq-mass-mat-m}
x=propagate(eq,151)
fbar(x)=FLQuant(c(seq(0,max(fbar(eq)),length.out=51),seq(1,100,length.out=101)*max(fbar(eq)))[-152],dimnames=list(iter=seq(151)))

@@

```{r, check1}
par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","ssb"]),k=0.1,w50=par["a"]*par["l50"]^par["b"]))

@@

om2=om

@@

catch(om2)=computeCatch(om2,slot="all")
plot(window(FLStocks("OM"=om,DD=om2),start=11))
```

**Figure `r iFig=iFig+1; iFig`** Check 1.

```{r, check2}
save(om2,file="~/Desktop/tmp/t.RData")
load("~/Desktop/tmp/t.RData")
ggplot(model.frame(FLQuants(om2[,dim(m(om2))[2]],"Biomass"=biomass,"Yield"=catch)))+
  geom_line(aes(Biomass,Yield),col="red")+
  geom_line(aes(Biomass,Yield),data=model.frame(FLQuants(om[,dim(m(om))[2]],"Biomass"=biomass,"Yield"=catch)))
```

**Figure `r iFig=iFig+1; iFig`** Check 2.

I think we need 4 OMs and 2 MPs

    OMs are
        OM with no DD,
        DD in mass,
        DD in mass + mat,
        DD in mass, mat + m
    MPs are HCR with
        refpts without DD (i.e. OM 1) and
        same as OM (for 2,3&4)

So we match/mismatch, in that we get the DD hypothesis right or use current practice

We condition all the OMs for M-at-age varying, and reported catch. An issue is the short-cut since if we assume no DD in the MP but there is DD in the OM we have to replicate the bias.

The remaining issue is random variation in recs, mass etc.
